### 2013-12-06 paper.tex
updated vals for signal

\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amssymb}
\newcommand{\ket}[1]{\ensuremath{\left|#1\right\rangle}}
\newcommand{\bra}[1]{\ensuremath{\left\langle#1\right|}}
\newcommand{\braket}[2]{\ensuremath{\left\langle #1 \middle| #2 \right\rangle}}
\newcommand{\obar}[1]{\ensuremath{\overline{ #1 }}}
\hypersetup{colorlinks=true, urlcolor=blue, linkcolor=blue, citecolor=red}
\graphicspath{ {C:/Users/Evan/Desktop/} }
\title{Relative Photometry of HD 227858 and HD 338931 from Landholt Standard SA 113475}
\author{Evan Ott \\ UT EID: eao466}
\date{\today}
\setcounter{secnumdepth}{0}
%\usepackage[parfill]{parskip}
\begin{document}
\maketitle
\section{Introduction}

\section{Methods}

\section{Results}

\begin{tabular}{c | c | c | c | c}
Star & Filter & $X=\sec{z}$ & $S_{*,sky}$ & $S_{sky}$ \\
\hline
HD 227858 & B & 1.147 & 549081 & 445607 \\
HD 227858 & I & 1.156 & 437742 &  182394\\
HD 227858 & R & 1.152 & 551039 & 438748\\
HD 227858 & V & 1.150 & 320125 & 282855\\
HD 338931 & B & 1.188 & 1078502 & 286905\\
HD 338931 & I & 1.197 & 982646 & 254194\\
HD 338931 & R & 1.194 & 996322 & 632289\\
HD 338931 & V & 1.191 & 526991 & 443543\\
SA 113475 & B & 1.150 & 139874 & 128880\\
SA 113475 & I & 1.153 & 461871 & 186586\\
SA 113475 & R & 1.151 & 643233 & 440040\\
SA 113475 & V & 1.151 & 367110 & 294031\\
SA 113475 & B & 1.428 & 565220 & 403841\\
SA 113475 & I & 1.475 & 1991118 & 1138438\\
SA 113475 & R & 1.454 & 251570 & 148624\\
SA 113475 & V & 1.441 & 318197 & 264164
\end{tabular}

\section{Analysis}
$S_{*,sky}=\Delta\nu{\obar{\Phi}}{\obar{f}}tA/g$

\section{Conclusion}

\begin{thebibliography}{9}

\end{thebibliography}

\end{document}