### 2013-12-10 paper.tex
COEFFICIENT

\documentclass{article}
\usepackage[latin1]{inputenc}
\usepackage{enumerate}
\usepackage{hyperref}
\usepackage{graphics}
\usepackage{graphicx}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tabularx}
\usepackage{amsmath}
\usepackage{amssymb}
\newcommand{\ket}[1]{\ensuremath{\left|#1\right\rangle}}
\newcommand{\bra}[1]{\ensuremath{\left\langle#1\right|}}
\newcommand{\braket}[2]{\ensuremath{\left\langle #1 \middle| #2 \right\rangle}}
\newcommand{\obar}[1]{\ensuremath{\overline{ #1 }}}
\hypersetup{colorlinks=true, urlcolor=blue, linkcolor=blue, citecolor=red}
\graphicspath{ {C:/Users/Evan/Desktop/} }
\title{Relative Photometry of HD 227858 and HD 338931 from Landholt Standard SA 113475}
\author{Evan Ott \\ UT EID: eao466}
\date{\today}
\setcounter{secnumdepth}{0}
%\usepackage[parfill]{parskip}
\begin{document}
\maketitle
\section{Introduction}

\section{Methods}

\section{Results}

\begin{center}
\begin{tabular}{c | c | c | c | c | c }
							%SUM              FLUX
Star & Filter & t(s) & $X=\sec{z}$ & $S_{*,sky}=\mathsf{SUM}$ & $S_*=\mathsf{FLUX=SUM-MSKY*AREA}$\\
\hline
HD 227858 & B & 30 & 1.294 & 117809 & 39381 \\
HD 227858 & V & 30 & 1.301 & 388133 & 131079  \\
HD 227858 & R & 30 & 1.307 & 567655 & 254168  \\
HD 227858 & I &  30 & 1.323 & 417430 & 260971   \\

HD 338931 & B & 30 & 1.394 & 255793 & 152570 \\
HD 338931 & V & 30 & 1.405 & 767017 & 494401  \\
HD 338931 & R & 30 & 1.413 & 1390626 &  977589 \\
HD 338931 & I & 30 &  1.424 & 1285304 &  1075136 \\

SA 113475 & B & 30 & 1.206  & 169993 & 38822  \\
SA 113475 & B & 30 & 1.482  & 179149 &  33471 \\

SA 113475 & V & 30 & 1.207  & 477466 & 154874 \\
SA 113475 & V & 30 & 1.498  & 665531 &  147737 \\

SA 113475 & R & 30 & 1.209 & 802783  & 344417  \\
SA 113475 & R & 30 & 1.498 & 1064716 & 332641 \\

SA 113475 & I & 30 & 1.213 & 477466  & 154874  \\
SA 113475 & I & 30 & 1.521& 665531 & 147737   \\

\end{tabular}
\end{center}

From \ref{eq:ext} and reference data for SA 113475, we arrive at the extinction coefficients:
\begin{align*}
k_B&= 0.583\\
k_V&= 0.176\\
k_R&= 0.131\\
k_I&= 0.166
\end{align*}
\begin{equation}
\label{eq:ext}
m_1-m_2=k(X_1-X_2)=-2.5\log_{10}\frac{S_1t_2}{S_2t_1}
\end{equation}
\section{Analysis}
$S_{*,sky}=\Delta\nu{\obar{\Phi}}{\obar{f}}tA/g$

\section{Conclusion}

\begin{thebibliography}{9}

\end{thebibliography}

\end{document}